Forecasting short-term hourly Emergency Departement arrivals

### 2022-01-06 paper/paper.Rmd
Draft conclusion

@choudhury2020forecasting used ARIMA, Holt-Winters, TBATS, and neutral network methods to forecast hourly accident and emergency department arrivals. ARIMA model was selected as the best fit model. Authors claimed that ARIMA has provided high and acceptable hourly ED forecasting accuracy, even outperforming TBATS.
<!-- @hertzum2017forecasting work was mentioned in this paper. This result is surprising given the existence of multiple seasonality in the hourly dataset. -->

@@

>>>>

@choudhury2020forecasting used ARIMA, Holt-Winters, TBATS, and neutral network methods to forecast hourly accident and emergency department arrivals. ARIMA model was selected as the best fit model. Authors claimed that ARIMA has provided high and acceptable hourly ED forecasting accuracy, even outperforming TBATS. @Cheng2021 develop an ARIMA model for ED occupancy with a seasonal component and exogenous variables, which is found to outperform a rolling-average benchmark. They also produce prediction intervals, a form of probabilistic forecast, which are found to be well calibrated, a necessary property for such forecasts. 
<!-- @hertzum2017forecasting work was mentioned in this paper. This result is surprising given the existence of multiple seasonality in the hourly dataset. -->

@@

Short-term forecasting of arrivals at emergency departments is an important element of staff and resource management in hospitals. Furthermore, due to the asymmetric impact of having an excess of shortage of resources, especially in emergency departments, quantifying forecast uncertainty is also of value as it enables planners to manage associated risks. Here, we have developed methods for producing probabilistic forecast of hourly arrivals up to 48 hours ahead, comparing different state-of-the-art approaches.

Two appraoches produced highly accurate, calibrated probabilistic forecasts, one time series method and one based on distributional regression. The first is ADAM-iESTX, which is an extension of exponential smoothing incorporating seasonality and assuming a Gamma predictive distribution. The second, labeled NOtr-2, regressed the two parameters of a truncated (at zero) normal distribution on features date and time features, and temperature. Both approaches produced calibrated probabilistic forecasts, but the point prediction produced by ADAM-iESTX produced forecasts had a lower RMSE than NOtr-2, and NOtr-2 produced forecasts with a lower pinball score. This suggests that the latter may be preferred if the whole distribution is used in decision-making.

Having compared the performance of a wide range of methods, we make the following observations: the choice of distribution assumed for probabilistic forecasts, and choice of model features, are as if not more important than the type of model employed; methods based on quantile regression, which do not assume a parametric distribution for forecasts, do not perform as well as those based on parametric distributions; and the best performing models handled the non-negative and skewed nature of the data automatically without need for post-processing. These observations reflect the characteristics of the data, which is representative of A\&E arrivals, but determining the extent to which they generalise is beyond the scope of this article. Furthermore, methods based on continuous valued distributions are not adversely affected by the fact that the data are interger-valued. Rounding up predictive quantiles to the next integer does not make predictions worse.

Finally, we have found that out-of-the-box models, those which require minimal tuning or manual development, do not perform as well as well-considered statistical methods. The popular TBATS, Prophet and Gradient Boosting Machine algorithms perform poorly compared to ADAM-iETSX and NOtr-2, and even the benchmarks. Of the models requiring a modest amount of user input and expertise, exponential smoothing (ETS) was found to perform well. ETS produces reasonably well calibrated forecasts, in contast to the benchmarks which were poorly calibrated, and highly accurate point forecasts, however, its probabilistic forecasts were considerably worse than  NOtr-2 in terms of pinball score.

Probabilistic forecasting opens the door to more sophisticated resource management in healthcare settings by providing decision-makers with uncertainty information and enabling quantitative risk management. Linking forecasts of arrivals with upstream (ambulance call-outs) and downstream (length of stay, medical outcomes) analytics offers an opportunity to improve forecast skill, and may also be necessary to maximise benefits though more holistic decision-making.
